---
title: 'Homework #5'
author: "Vitor Borges"
date: "08/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Working with matrices, lists and data frames

1. Assign to the variable `n_dims` a single random integer between 3 and 10.

```{r, eval = TRUE, echo = TRUE}

n_dims <- sample(x = 3:10, size = 1)

```
- Create a vector of consecutive integers from 1 to _n_dims²_.

```{r, eval = TRUE, echo = TRUE}

vec1 <- 1:n_dims^2

```

- Use the sample function to randomly reshuffle these values.

```{r, eval = TRUE, echo = TRUE}

vec2 <- sample(x = vec1, size = length(vec1), replace = FALSE)


```

- create a square matrix with these elements.

```{r, eval = TRUE, echo = TRUE}

sq_mat <- matrix(data = vec2, nrow = 9, ncol = 9)

```

- print out the matrix.

```{r, eval = TRUE, echo = TRUE}

print(sq_mat)

```

- find a function in r to transpose the matrix.

```{r, eval = TRUE, echo = TRUE}

tsq_mat <- t(sq_mat)

```

- print it out again and note how it has changed.

```{r, eval = TRUE, echo = TRUE}

print(tsq_mat)

```

- calculate the sum and the mean of the elements in the first row and the last row.

```{r, eval = TRUE, echo = TRUE}

# sum and meam of first and last rows using functions sum() and mean()

first_rowsum <- sum(sq_mat[1, ])
last_rowsum <- sum(sq_mat[nrow(sq_mat), ])

first_rowmean <- mean(sq_mat[1, ])
last_rowmean <- mean(sq_mat[nrow(sq_mat), ])

# sum and meam of first and last rows using functions rowSums() and rowMeans()

first_last_rowsum <- rowSums(sq_mat)[c(1, nrow(sq_mat))]
first_last_rowmean <- rowMeans(sq_mat)[c(1, nrow(sq_mat))]

print(cat("sum of first row =", first_last_rowsum[1], "and mean of first row =", first_last_rowmean[2]))
print(cat("sum of last row =", first_last_rowsum[2], "and mean of last row =", first_last_rowmean[2]))

```

- read about the `eigen()` function and use it on your matrix

```{r, eval = TRUE, echo = TRUE}

spec_decomp <- eigen(x = sq_mat)

```

- look carefully at the elements of $values and $vectors. What kind of numbers are these?

- dig in with the typeof() function to figure out their type.

```{r, eval = TRUE, echo = TRUE}

typeof(spec_decomp$values) 
typeof(spec_decomp$vectors) 

```

- if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.

2. Create a list with the following named elements:

- `my_matrix`, which is a 4 x 4 matrix filled with random uniform values
- `my_logical` which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
- `my_letters`, which is a 26-element vector of all the lower-case letters in random order.

```{r, eval = TRUE, echo = TRUE}

my_matrix <- matrix(data = runif(n = 16), nrow = 4, ncol = 4)
my_logical <- runif(n = 100) > 0.6
my_letter <- sample(x = letters, replace = FALSE)
my_list <- list(my_matrix, my_logical, my_letter)

```

- create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
- use the typeof() function to confirm the underlying data types of each component in this list
- combine the underlying elements from the new list into a single atomic vector with the c() function.
- what is the data type of this vector?

```{r, eval = TRUE, echo = TRUE}

new_list <- list(my_list[[1]][2, 2], my_list[[2]][2], my_list[[3]][2])

typeof(new_list[[1]])
typeof(new_list[[2]])
typeof(new_list[[3]])

new_vector <- c(new_list[[1]], new_list[[2]], new_list[[3]])

typeof(new_vector)

```

3. Create a data frame with two variables (= columns) and 26 cases (= rows).

- call the first variable my_unis and fill it with a random uniform value from 0 to 10

- call the second variable my_letters and fill it with 26 capital letters in random order.

```{r, eval = TRUE, echo = TRUE}

data <- data.frame(my_unis = runif(n = 26, min = 0, max = 10), my_letters = sample(x = LETTERS, replace = FALSE))

```

- for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.

```{r, eval = TRUE, echo = TRUE}

data[sample(nrow(data), size = 4, replace = FALSE), 1] <- NA
data

```

- for the first variable, write a single line of R code to identify which rows have the missing values.

```{r, eval = TRUE, echo = TRUE}

which(is.na(data$my_unis))
which(!complete.cases(data$my_unis))

```

- for the second variable, sort it in alphabetical order

```{r, eval = TRUE, echo = TRUE}

sort(data$my_letters, decreasing = FALSE)

```

- calculate the column mean for the first variable.

```{r, eval = TRUE, echo = TRUE}

mean(x = data$my_unis, na.rm = TRUE)

```
